---
title: "USBusinessReport"
author: "Simone Trindade Steel"
date: "20/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## US Business Report: an analysis of economic sentiment and new housing construction

This reports is based on a study of business and industry trends in the United States 
using the US Census Bureau data.
The inspiration for this study is the hypothesis that US National level of value generation,
as an indication of economic optimism, drives the antecipation of new property demand. 

The analysis looks for regional variations in the US and attempts to predict the regional 
growth of new home construction and sales.

The source of data is https://www.kaggle.com/census/business-and-industry-reports



```{r set-up, include=FALSE}
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(tidytext)
library(caret)
library(data.table)
library(rpart)
library(randomForest)


options(digits = 3)    # 3 significant digits

getwd()
filename <- "CapstoneProject_data.csv"
metadata <- "CapstoneProject_metadata.csv"

# Loading time series data
temp <- read_file(filename)
data <-read_csv(temp, guess_max = 1000000)
df <- as.data.frame(data)
df <- df %>% mutate(value = as.numeric(value))


# Loading metadata to enable selection of Housing indicators and overall Financial results
temp <- read_file(metadata)
md <-read_csv(temp)
md_df <- as.data.frame(md)

```

## Understanding the data

The original dataset has many economic factors that will not be used in this study.
The time series that are relevant to the subject of this report are the 
macroeconomic indicators (Financial Reports) and the new housing indicators 
(New Home Sales, New Residential Construction)

All dates indicate the begining of the analysis period, and they have been used to align 
the time series into quarterly periods.

The graphs below summarise the US national totals, and the disruption created by the 2008 crisis. Following the crisis, revenue resumed its upwards trajectory, as seen with broader economic recovery. The analysis
will also consider variations of revenue generation per industry, as these also exhibit geographic characteristics.

The regional graphs, however, show that new housing construction and offering vary along each individual regional average that appears not related to the broader revenue generation.

In order to understand what drives an increase and decrease of new housing, 
this analysis will classify trend as positive when the numbers are above average, and as negative 
when they are below in each distinct region.


```{r analysis, include=FALSE}
# What time series are used for new home sales and housing under construction indicators? 
housing_cat <- md_df %>% 
  filter((report=="New Residential Construction" | report=="New Home Sales"), 
         dt_code=="TOTAL",
         is_adj==0)
str(housing_cat)

# What time series are used for value creation in the financial reports (sales, revenue)? 
finance_cat <- md_df %>% 
  filter(report=="Quarterly Financial Report", 
         dt_code=="101",
         is_adj==0)
str(finance_cat)



# Loading the relevant time series for analysis

# All activity related to new dwellings, as a proxy for expected regional demand for housing
housing_dat <- df %>% filter(time_series_code %in% housing_cat$time_series_code)
head(housing_dat)

# All sales and revenue at national level, keeping all the aggregated numbers per industry (category_level=0)
finance_cat <- finance_cat %>% filter(category_level==0)
finance_dat <- df %>% filter(time_series_code %in% finance_cat$time_series_code)
# Separate industry category as a column
finance_dat <- finance_dat %>% mutate(industry = str_trunc(time_series_code, 3, side="right", ellipsis=""))
finance_dat <- finance_dat %>% mutate(industry = as.factor(industry))
head(finance_dat)

# Trimming the housing time series according to available financial data
housing_dat <- housing_dat %>% filter(date>=min(finance_dat$date) & date<=max(finance_dat$date))
min(housing_dat$date) == min(finance_dat$date)
max(housing_dat$date) == max(finance_dat$date)


# Aggregate quarterly for analysis alongside financial data
housing_dat <- housing_dat %>% 
  mutate(quarter = as_date(ifelse(month(date)==2 | month(date)==3,
                                               as_date(paste(as.character(year(date)),'01',as.character(day(date)), sep="-")),
                           ifelse(month(date)==5 | month(date)==6,
                                               as_date(paste(as.character(year(date)),'04',as.character(day(date)), sep="-")),
                           ifelse(month(date)==8 | month(date)==9,
                                               as_date(paste(as.character(year(date)),'07',as.character(day(date)), sep="-")),
                           ifelse(month(date)==11 | month(date)==12,
                                               as_date(paste(as.character(year(date)),'10',as.character(day(date)), sep="-")),
                                               date))))))

housing_dat %>% group_by(quarter) %>% summarise(n())

# Override date column for easier manipulation later
housing_dat <- housing_dat %>% mutate(date=quarter)

# Separate state and national levels
housing_dat <- housing_dat %>% mutate(region = str_trunc(time_series_code, 2, side="left", ellipsis=""))
housing_dat <- housing_dat %>% mutate(region = as.factor(region))
# Aggregating all new house activity
housing_dat <- housing_dat %>% group_by(date, region) %>% summarise(value=sum(value))
housing_dat_total <- data.frame(housing_dat)
str(housing_dat_total)

# Visualising the effects of National value generation on the antecipation of housing demand (i.e. new build) regionally.

```

```{r analysis-graphs, echo=FALSE}
# Firstly, the 2008 housing market shock is clearly visible in both the financial and housing data
graph_fin <- finance_dat %>% group_by(date) %>% summarise(value_generated=sum(value)) %>% 
  ggplot(aes(x=date, y=value_generated)) + 
  ylab("Sales and Revenue in USD Millions") +
  ggtitle("US total from Q4/2000 to Q1/2017") +
  geom_line() + geom_vline(xintercept=date("2008-09-15"), colour="red") +
  annotation_custom(grid::linesGrob())
graph_fin

avg <- housing_dat %>% ungroup() %>% filter(region=="US") %>% select(v=value) %>% summarise(mean(v))
graph_h_us <- housing_dat_total %>% filter(region=="US") %>% 
   group_by(region, date) %>% summarise(new_builds=sum(value)) %>% 
   ggplot(aes(x=date, y=new_builds)) + 
   ylab("Thousands of new housing units") +
   ggtitle("US total from Q4/2000 to Q1/2017") +
   geom_point() + geom_hline(yintercept=as.numeric(avg), colour="blue") +
   geom_vline(xintercept=date("2008-09-15"), colour="red")
graph_h_us

# Looking closer at each region, the volatility in the South is far greater than the other regions.
# The Northeast is stable relatively to the national income fluctuations.
graph_h_reg <- housing_dat_total %>% filter(region!="US") %>% 
  group_by(region, date) %>% summarise(new_builds=sum(value)) %>% 
  ggplot(aes(x=date, y=new_builds, colour=region)) + 
  ylab("Thousands of new housing units") +
  ggtitle("Total per region from Q4/2000 to Q1/2017") +
  geom_smooth() 
graph_h_reg


```


## Preparing data for analysis

Several steps are performed here:

(1) Aligning the time periods between income generation and house construction, allowing for a 1 year lag between the financial indicators being perceived and the construction to yield new houses.

(2) Removing aggregated US National data points in order not to double count properties.

(3) Preparing regional averages, given that the variations are significant between regions.

(4) Introducing industry classification, given their regional variation.

Industry codes are:


```{r preparing-data, echo=FALSE}

# Industry codes and description
finance_cat %>% select(Code=cat_code, Description=cat_desc) %>% knitr::kable()

# Adding a lag of 1 year based on sentiment of growth (sales, revenue) and new house activity
housing_dat_total <- housing_dat_total %>% mutate(previous_date = as_date(paste(as.character(year(date)-1),
                                                   as.character(month(date)),
                                                   as.character(day(date)), sep="-")))
housing_dat_total <- housing_dat_total %>% mutate(date=previous_date)
#head(housing_dat_total)

# Matrix for prediction, aligning all values by quarter
dataset <- inner_join(finance_dat, housing_dat_total, by = "date")
dataset <- dataset %>% mutate(new_houses_K = value.y, 
                              national_income_M = value.x) %>%
                       dplyr::select(date, region, industry, new_houses_K, national_income_M)

#head(dataset)

# Removing US totals
dataset <- dataset %>% filter(!is.na(national_income_M), region!="US")

# Pre 2008 (darker shades), construction was higher in the South and West regions.
# But over time, it seems negatively influenced by income, especially in the Mining industry
dataset %>% 
  ggplot(aes(x=log10(national_income_M), y=new_houses_K, colour=year(date))) + 
  xlab("Sales and Revenue in USD Millions (log10)") +
  ylab("New Housing in Thousands of Units") +
  ggtitle("Revenue and New Housing, per region and industry") +
  geom_point()  + 
  facet_grid(industry ~ region)

# This requires centreing the analysis on regional means

housing_avg <- dataset %>% group_by(region) %>% summarise(avg = mean(new_houses_K))
dataset <- inner_join(dataset, housing_avg, by="region")

# Trend = 1 meaning positive, if more new houses than average are predicted
# Trend = 0 meaning negative, if fewer new houses than average are predicted

dataset <- dataset %>% mutate(trend=as.factor(ifelse(new_houses_K>=avg,1,0)))
#head(dataset)

```


## Making predictions

This report aims to demonstrate that the regional trends of new house construction is lagging 
the sentiment of a positive economic position, as observed by financial reports on sales, invoicing and revenue at the national level.

In order to do that, the concept of positive or negative trend was introduced. Positive trend means
that regional construction is above historical average, and negative trend means below average.

The models that appeared most appropriate for this exercise were GLM, KNN and Random Forest.

The ensemble of those models did not improve on the prediction made via Random Forest.

```{r prediction, include=FALSE}

# Splitting training and test datasets

# Validation set will be 50% of Housing data
set.seed(1, sample.kind="Rounding")

test_index <- createDataPartition(y = dataset$new_houses_K, times = 1, p = 0.5, list = FALSE)
train_ds <- dataset[-test_index[,1],]
test_ds <- dataset[test_index[,1],]



# Training 

# Selecting predictors (Training Set)
x <- train_ds %>% dplyr::select (date, region, industry, national_income_M, avg)
y <- train_ds$trend


# ============= GLM ==============
fit <- train(x, y, method = "glm")
fit$results
# Testing
y_hat_GLM <- predict(fit, test_ds) 
y_hat <- y_hat_GLM
acc <- confusionMatrix(data = y_hat, reference = test_ds$trend)$overall["Accuracy"]
res <- data.frame(model="GLM", acc=acc)


# ========= Random Forest ========
fit <- randomForest(y ~ ., data = x, nodesize = 50, maxnodes = 25)
fit$importance
# Testing
y_hat_RF <- predict(fit, test_ds) 
y_hat <- y_hat_RF
acc <- confusionMatrix(data = y_hat, reference = test_ds$trend)$overall["Accuracy"]
res <- rbind(res, data.frame(model="RF", acc=acc))

# ========= KNN ============

ks <- seq(1, 31, 3)
F_1 <- sapply(ks, function(k){
  fit <- knn3(y ~ ., data = x, k = k)
  y_hat <- predict(fit, test_ds, type = "class") %>% 
    factor(levels = levels(test_ds$trend))
  F_meas(data = y_hat, reference = test_ds$trend)
})
# plot(ks, F_1)
# max(F_1)
# ks[which.max(F_1)]
fit <- knn3(y ~ ., data = x, k = ks[which.max(F_1)])
y_hat_KNN <- predict(fit, test_ds, type = "class")
y_hat <- y_hat_KNN
acc <- confusionMatrix(data = y_hat, reference = test_ds$trend)$overall["Accuracy"]
res <- rbind(res, data.frame(model="KNN", acc=acc))


# ======= Ensemble ===========
# ENSEMBLE #
y_set <- data.frame(y_hat_GLM = as.integer(ifelse(y_hat_GLM==1, 1, 0)),
                    y_hat_KNN = as.integer(ifelse(y_hat_KNN==1, 1, 0)),
                    y_hat_RF = as.integer(ifelse(y_hat_RF==1, 1, 0)))

y_set <- y_set %>% mutate(ens_y_hat = round((y_hat_GLM + y_hat_KNN + y_hat_RF)/3))
                                               
y_set <- y_set %>% mutate(y_hat = as.factor(ens_y_hat))
levels(y_set$y_hat) <- levels(test_ds$trend)
#head(y_set)
acc <- confusionMatrix(y_set$y_hat, test_ds$trend)$overall["Accuracy"]
res <- rbind(res, data.frame(model="Ensemble", acc=acc))

```

```{r diplay-predictions, echo=FALSE}
res %>% select(Model=model, Accuracy=acc) %>% knitr::kable()

```



## Conclusion

There is a reasonably high predictive power of house construction by US Region based on the past year financial performance of each industry. The maximum accuracy obtained with the Random Forest model was 0.959.

Armed with this knowledge, there is potential for better decision making and reduction of unintended, excessive housing supply.

